---
title: "yarrr package guide"
author: "Nathaniel Phillips (yarrr.book@gmail.com)"
date: "`r Sys.Date()`"
csl: apa.csl
bibliography: yarrr.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# YaRrr! The pirate's guide to R

This package contains data, functions, and other random files for the e-book YaRrr! The pirate's guide to R. You can access the latest version of the book at [www.thepiratesguidetor.com](www.thepiratesguidetor.com).


# Guides

To see guides for the main functions, click one of the following links:

- [pirateplot() - A function for creating "Pirate Plots"](pirateplot.html)
- [piratepal() - A function to create pirate-y color palettes](piratepal.html)

# Examples

## pirateplot()

The pirateplot() function creates a pirateplot

```{r, echo = F, message = F, results = 'hide'}
library(yarrr)
```

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
pirateplot(formula = weight ~ Time,
           data = ChickWeight,
           main = "Chicken weights by Time (week)")
```

## piratepal()

The `piratepal()` function returns different color palettes either created by graphic designers, or inspired by random things I have found around my office `piratepal("espresso")`, or horror movies `piratepal("evildead)"`. Here are all of the palettes:

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}
piratepal("all")
```

You can then use specific palettes in your plots by specifying the named palette. I'll create a scatterplot using the google palette with `piratepal('google', trans = .5)`:

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}

my.cols <- piratepal(palette = "google",  
                     trans = .5)

set.seed(100) # For reproducibility
x <- rnorm(100)
y <- x + rnorm(100)

plot(x = x, y = y, col =  my.cols,
     pch = 16,
     cex = runif(100, min = 0, max = 2),
     main = "piratepal('google', trans = .5)")
```


# Questions

If you have any questions, comments, or suggestions, write me at yarrr.book@gmail.com

### References
