---
title: "piratepal(): Color palettes for R Pirates"
output: rmarkdown::html_vignette
bibliography: yarrr.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = F, message = F, results = 'hide'}
library("yarrr")
```

# piratepal()

`piratepal()` is a function that returns color palettes. The function has three main arguments:

- `palette`: The specific palette (entered as a string) you want to use. You can also specify `palette = "all"` to see all the palettes, or `palette = "random"` to get a random palette.

- `trans`: A number between 0 and 1 indicating how transparent to make the colors.

- `length.out`: An optional number indicating how many colors to return (if `length.out` is larger than the number of colors in the palette, they will just be repeated).

- `plot.result`: A logical value indicating whether or not to display the palette.

# Examples

To see all of the palettes, run `piratepal("all")`

```{r, fig.width = 6, fig.height = 6, echo = F, fig.align='center'}
piratepal(palette = "all")
```

Once you find a palette you'd like to use, you can return the colors as a vector by specifying the name of the palette in the `palette` argument:

```{r}
piratepal(palette = "ipod")
```

If you want to see a palette in detail (and possibly some images that inspired the palette), include the argument `plot.result = T`:

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
piratepal(palette = "southpark", 
          plot.result = T)
```

Some of the palettes are darker than others...

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
piratepal(palette = "evildead", 
          plot.result = T)
```


You can use palettes in your plots by entering the palette as color arguments. I'll create a scatterplot using the google palette with `piratepal('nemo', trans = .5)`:

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}

nemo.cols <- piratepal(palette = "nemo",  
                     trans = .5)

set.seed(100) # For reproducibility
x <- rnorm(100)
y <- x + rnorm(100)

plot(x = x, y = y, col =  nemo.cols,
     pch = 16,
     cex = runif(100, min = 0, max = 2),
     main = "piratepal('nemo', trans = .5)")
    
```


## Plotting the Up house

Let's use the basel palette to draw the house from Up.

```{r, fig.width = 6, fig.height = 6, fig.align='center'}

# Plot the Up house using colors from the "google" palette

balloon.colors <- piratepal("basel", trans = .1)
n.balloons <- 500
balloon.x <- rnorm(n.balloons, 0)
balloon.y <- rnorm(n.balloons, 4, 1)

plot(1, xlim = c(-15, 7), ylim = c(-15, 7),
    xlab = "", ylab = "", type = "n", 
    xaxt = "n", yaxt = "n", bty = "n")

# house
rect(-2, -6, 2, -2)
polygon(c(-2, 0, 2),
        c(-2, 0, -2))
rect(-.5, -6, .5, -4)
points(.3, -5)

line.start.x <- rnorm(n.balloons, 0, .2)
line.start.y <- -1 + rnorm(n.balloons, 0, .1)

# strings
segments(line.start.x, 
         line.start.y, 
         balloon.x, balloon.y, 
         lty = 1, col = gray(.5, .1), lwd = .2)

# balloons
points(balloon.x, balloon.y, pch = 21, 
       bg = balloon.colors,
       col = gray(.9), cex = rnorm(100, 2, .3))
```



# References


