---
title: "pirateplot() guide"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: yarrr.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# What is a pirateplot()?

A pirateplot() is the RDI plotting choice of R pirates who are displaying the relationship between 1 or two categorical variables, and a continuous dependent variable. A pirateplot() has the option to display Raw data (as individual points), Descriptive statistics (i.e.; group means as bars or lines), and Inferential statistics (i.e.; Bayesian Highest Density Intervals or Confidence intervals) in the same plot.

A pirateplot() has 4 distinct elements

1. points, symbols representing the raw data (jittered horizontally)
2. bar, a vertical bar showing central tendencies
3. bean, a smoothed density (inspired by @kampstra2008beanplot) representing a smoothed density
4. box, a rectangle representing an inference interval (either a Bayesian Highest Density Interval or a frequentist confidence interval)

```{r, fig.width = 6, fig.height = 6, echo = F, fig.align='center'}
yarrr::pirateplot(formula = weight ~ Diet,
                  data = ChickWeight,
                  theme.o = 1,
                  bean.o = c(0, .1, 1, .1),
                  point.o = c(.4, .1, .1, .1),
                  line.o = c(.5, .5, .5, .5),
                  inf.o = c(.1, .1, .1, .9),
                  bar.o = c(.1, .8, .1, .1),
                  main = "4 Elements of a pirateplot()"
                  )


text(.7, 350, labels = "Points")
text(.7, 350, labels = "Raw Data", pos = 1, cex = .8)
arrows(.7, 320, .97, 270, length = .1)

text(1.35, 200, labels = "Bar")
text(1.35, 200, labels = "Center", pos = 1, cex = .8)
arrows(1.35, 170, 1.54, 125, length = .1)

text(2.4, 250, labels = "Bean")
text(2.4, 250, labels = "Density", pos = 1, cex = .8)
arrows(2.4, 220, 2.85, 200, length = .1)

text(3.4, 300, labels = "4\nBox")
arrows(3.4, 270, 3.8, 150, length = .1)


```





# Examples

Here is a pirateplot() of the weights of chickens from the ChickWeight dataset:

```{r, fig.width = 6, fig.height = 6}
yarrr::pirateplot(formula = weight ~ Diet,
                  data = ChickWeight,
                  theme.o = 3)
```





# References


