---
title: "piratepal(): Color palettes for R Pirates"
output: rmarkdown::html_vignette
bibliography: yarrr.bib
vignette: >
  %\VignetteIndexEntry{Pirate-y color palettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = F, message = F, results = 'hide'}
library("yarrr")
```

## piratepal()

`piratepal()` is a function that returns color palettes. The function has four main arguments:

- `palette`: The specific palette you want to use. For example, `palette = "appletv"` will give you the appletv palette. You can also specify `palette = "all"` to see all the palettes, or `palette = "random"` to get a random palette.

- `trans`: A number between 0 and 1 indicating how transparent to make the colors. 1 is fully transparent (e.g. invisible), and 0 is not transparent at all. Personally, I like at least slightly transparent colors (e.g.; `trans = .3`)

- `length.out`: An optional number indicating how many colors to return (if `length.out` is larger than the number of colors in the palette, they will just be repeated).

- `plot.result`: A logical value indicating whether or not to display the palette.

## Examples

To see all of the palettes, run `piratepal("all")`

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
piratepal(palette = "all")
```

Once you find a palette you'd like to use, you can return the colors as a vector by specifying the name of the palette in the `palette` argument. Here is the Google palette

```{r}
piratepal(palette = "google")
```

If you want to see a palette in detail (and possibly some images that inspired the palette), include the argument `plot.result = T`. Here's where the `southpark` palette comes from.

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
piratepal(palette = "southpark", 
          trans = .5,
          plot.result = T)
```


You can look at a random palette by setting `palette = "random"`:

```{r echo = F}
set.seed(105)
```


```{r, fig.width = 6, fig.height = 6, fig.align='center'}
piratepal("random", plot.result = T)
```



Some of the palettes are darker than others. Here's the nightmare-inducing evildead palette (I'll up the transparency to .5 to make it a bit less scarring...)

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
piratepal(palette = "evildead",
          trans = .5,
          plot.result = T)
```

Here's a scatterplot using the xmen palette with `piratepal('xmen', trans = .5)`:

```{r, fig.width = 6, fig.height = 6, fig.align = 'center'}
xmen.cols <- piratepal(palette = "xmen",  
                     trans = .5)

x <- rnorm(100)
y <- x + rnorm(100)

plot(x = x, y = y, 
     col =  xmen.cols,
     pch = 16,
     cex = 2,
     main = "piratepal('xmen', trans = .5)")
```


### Plotting the Up house

Let's use the basel palette to draw the house from Up [Up Trailer](https://www.youtube.com/watch?v=pkqzFUhGPJg)

```{r, fig.width = 8, fig.height = 6, fig.align='center', results='hold'}
# Set up balloons
balloon.colors <- piratepal("basel", trans = .2)
balloon.x <- rnorm(500, 0)
balloon.y <- rnorm(500, 4, 1)

par(mar = rep(.1, 4))
plot(1, xlim = c(-15, 7), ylim = c(-15, 7),
    xlab = "", ylab = "", type = "n", 
    xaxt = "n", yaxt = "n", bty = "n")

# skyline
start.x <- runif(200, -15, 7)
start.y <- sort(runif(200, -15, -12), decreasing = T)
heights <- runif(200, 2, 4)
widths <- runif(200, .25, 1.5)
  
rect(start.x, start.y, start.x + widths, start.y + heights, 
     col = "white", border = gray(.4))

# house
rect(-2, -6, 2, -2)
polygon(c(-2, 0, 2),
        c(-2, 0, -2))
rect(-.5, -6, .5, -4)
points(.3, -5)

# strings
line.start.x <- rnorm(500, 0, .2)
line.start.y <- -1 + rnorm(500, 0, .1)
segments(line.start.x, 
         line.start.y, 
         balloon.x, balloon.y, 
         lty = 1, col = gray(.5, .1), lwd = .2)

# balloons
points(balloon.x, balloon.y, pch = 21, 
       bg = balloon.colors,
       col = gray(.9), cex = rnorm(100, 2, .3))
```

```{r, echo = F}
par(mar = c(5, 4, 4, 1) + .1)
```


### Have a favorite palette?

If you have a favorite palette that you'd like me to add, just contact me at yarrr.book@gmail.com and I'll see what I can do :)
