---
title: "pirateplot() guide in the yarrr package"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: yarrr.bib
vignette: >
  %\VignetteIndexEntry{Pirateplot: The plotting choice of R pirates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# What is a pirateplot()?

A pirateplot, from the `yarrr` package (`devtools::install_github('ndphillips/yarrr', build_vignette = T`) is the RDI (**Raw** data, **Descriptive** statistics, and **Inferential** statistics) plotting choice of R pirates who are displaying the relationship between 1 or two categorical independent variables, and one continuous dependent variable.

```{r, echo = F, message = F, result = 'hide'}
library(yarrr)
```


A pirateplot() has 4 distinct elements

1. points, symbols representing the raw data (jittered horizontally)
2. bar, a vertical bar showing central tendencies
3. bean, a smoothed density (inspired by @kampstra2008beanplot) representing a smoothed density
4. inf, a rectangle representing an inference interval (either a Bayesian Highest Density Interval or a frequentist confidence interval)

```{r, fig.width = 6, fig.height = 6, echo = F, fig.align='center'}
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           theme.o = 1,
           bean.o = c(0, .1, 1, .1),
           point.o = c(.4, .1, .1, .1),
           line.o = c(.5, .5, .5, .5),
           inf.o = c(.1, .1, .1, .9),
           bar.o = c(.1, .8, .1, .1),
           main = "4 Elements of a pirateplot()", 
           pal = "google"
          )


text(.7, 350, labels = "Points")
text(.7, 350, labels = "Raw Data", pos = 1, cex = .8)
arrows(.7, 320, .97, 270, length = .1)

text(1.35, 200, labels = "Bar")
text(1.35, 200, labels = "Center", pos = 1, cex = .8)
arrows(1.35, 170, 1.54, 125, length = .1)

text(2.4, 250, labels = "Bean")
text(2.4, 250, labels = "Density", pos = 1, cex = .8)
arrows(2.4, 220, 2.85, 200, length = .1)

text(3.4, 300, labels = "Inf")
text(3.4, 300, labels = "Inference", pos = 1, cex = .8)

arrows(3.4, 270, 3.8, 150, length = .1)
```

```{r, echo = F, message = F, results='hide'}
library(yarrr)
```


# Main arguments

The major arguments to pirateplot() are a dataframe and a formula in the form `formula = y ~ x1 + x2`, where `y` is a dependent variable, `x1` is a categorical independent variable, and `x2` is an optional second independent variable. You can also include most of the the main plotting parameters that you would use for normal plotting functions like `main()`, `xlab()`, `ylim()` (etc.)

```{r, fig.width = 6, fig.height = 5, fig.align='center'}
pirateplot(formula = time ~ cleaner + type,
           data = poopdeck,
           main = "Poopdeck Cleaning Time"
           )
```


# Colors

You can adjust colors in many ways. You can select an overall color palette using the pal argument, where pal is the name of a color palette in the piratepal() function (to see all the palettes, run `piratepal(palette = "all", action = "s")`)

```{r, fig.width = 6, fig.height = 6}
pirateplot(formula = weight ~ Time,
           data = ChickWeight,
           pal = "southpark"   # Use the southpark color palette
           )
```

If you specify a single color name in the `pal` argument, all plotting elements will be that color:

```{r, fig.width = 6, fig.height = 5, fig.align='center'}
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           pal = "black"
           )
```

You can override the color of any specific element by using one of the x.col arguments (e.g.; `bar.col`, `point.col`, `inf.col`, `line.col`). In this example, I'll change the palette to `southpark` (try running `yarrr::piratepal("southpark", "show")`), and then make the points and bars gray.

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
pirateplot(formula = weight ~ Time,
           data = ChickWeight,
           pal = "southpark",   # Use the southpark color palette,
           point.col = "gray",  # black points
           bar.col = "gray"   # gray bars
           )
```




# Opacities

You can adjust the opacities of individual plotting elements in one of two ways. The `theme.o` argument allows you to specify an opacity theme. There are currently 3 themes supported. For example, `theme.o = 2` will turn off the inference boxes and turn up the bars:

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           theme.o = 2, # Use opacity theme 2
           main = "Opacity themes with theme.o"
           )
```

Or, you can turn off all of the opacities with `theme.o = 0` and then selectively turn up individual elements with `bar.o`, `bean.o`, `point.o`, and `inf.o`:

```{r, fig.width = 6, fig.height = 6, eval = F, echo = F, fig.align='center'}
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           theme.o = 0, # Turn off all opacities,
           point.o = .1, # Turn point opacities up to .1
           inf.o = .3, # Turn inf opacities up to .3
           bean.o = .9, # Turn beans way up
           main = "Changing opacities individually with point.o, inf.o (etc.)"
           )
```

Here is a plot with only the points
```{r, fig.width = 6, fig.height = 6, fig.align='center'}
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           theme.o = 0, # Turn off all opacities,
           point.o = .1 # Turn point opacities up to .1
           )
```

Now I'll add the beans
```{r, fig.width = 6, fig.height = 6, fig.align='center'}
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           theme.o = 0, 
           point.o = .1,
           bean.o = .5 # Turn on beans to .5
           )
```

Or just make a barplot with HDIs (and a gray palette)

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
pirateplot(formula = weight ~ Time,
           data = ChickWeight,
           theme.o = 0, 
           bar.o = .5, # Turn bar up to .5
           inf.o = .8, # Turn 95% HDI to .8
           pal = gray(.1)
           )
```


You can specify the x.o arguments as *vectors* to selectively turn up (or down) specific elements for specific groups:

```{r, fig.width = 6, fig.height = 6, fig.align='center'}
pirateplot(formula = weight ~ Diet,
           data = ChickWeight,
           pal = "basel",
           theme.o = 0, # Turn off all opacities,
           point.o = c(.1, .05, .2, .05), # Selectively turn up point opacities
           inf.o = c(.9, .1, .9, .1), # Selectively turn up inf opacities
           bean.o = c(1, .1, 1, .1), # Selective bean opacity adjustment
           main = "Adjusting opacities between groups"
           )
```

# Additional arguments

You can use up to two independent variables in the formula, change the color of the background with `back.col`, and the color of the gridlines with `gl.col`

```{r, fig.width = 10, fig.height = 6}
pirateplot(formula = weight ~ Diet + Time,
           data = ChickWeight,
           pal = "google",
           theme.o = 1, # Turn off all opacities,
           back.col = gray(.96),
           gl.col = "white",
           main = "back.col, gl.col"
           )
```

# References


