---
title: "pirateplot() guide"
author: "Nathaniel Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: yarrr.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# What is a pirateplot()?

A pirateplot() is the RDI plotting choice of R pirates who are displaying the relationship between 1 or two categorical variables, and a continuous dependent variable. A pirateplot() has the option to display Raw data (as individual points), Descriptive statistics (i.e.; group means as bars or lines), and Inferential statistics (i.e.; Bayesian Highest Density Intervals or Confidence intervals) in the same plot.

A pirateplot() has 4 distinct elements

1. points, symbols representing the raw data (jittered horizontally)
2. bar, a vertical bar showing central tendencies
3. bean, a smoothed density (inspired by @kampstra2008beanplot) representing a smoothed density
4. box, a rectangle representing an inference interval (either a Bayesian Highest Density Interval or a frequentist confidence interval)

```{r, fig.width = 6, fig.height = 6, echo = F, fig.align='center'}
yarrr::pirateplot(formula = weight ~ Diet,
                  data = ChickWeight,
                  theme.o = 1,
                  bean.o = c(0, .1, 1, .1),
                  point.o = c(.4, .1, .1, .1),
                  line.o = c(.5, .5, .5, .5),
                  inf.o = c(.1, .1, .1, .9),
                  bar.o = c(.1, .8, .1, .1),
                  main = "4 Elements of a pirateplot()"
                  )


text(.7, 350, labels = "Points")
text(.7, 350, labels = "Raw Data", pos = 1, cex = .8)
arrows(.7, 320, .97, 270, length = .1)

text(1.35, 200, labels = "Bar")
text(1.35, 200, labels = "Center", pos = 1, cex = .8)
arrows(1.35, 170, 1.54, 125, length = .1)

text(2.4, 250, labels = "Bean")
text(2.4, 250, labels = "Density", pos = 1, cex = .8)
arrows(2.4, 220, 2.85, 200, length = .1)

text(3.4, 300, labels = "Box")
text(3.4, 300, labels = "Inference", pos = 1, cex = .8)

arrows(3.4, 270, 3.8, 150, length = .1)


```



# Colors

You can adjust colors in many ways. You can select an overall color palette using the pal argument, where pal is the name of a color palette in the piratepal() function (to see all the palettes, run piratepal(palette = "all", action = "s"))

```{r, fig.width = 6, fig.height = 6}
yarrr::pirateplot(formula = weight ~ Time,
                  data = ChickWeight,
                  pal = "southpark"   # Use the southpark color palette
                  )
```

You can override the color of any specific element by using one of the x.col arguments (e.g.; bar.col, point.col, inf.col, line.col). In this example, I'll change

```{r, fig.width = 6, fig.height = 6}
yarrr::pirateplot(formula = weight ~ Time,
                  data = ChickWeight,
                  pal = "southpark",   # Use the southpark color palette,
                  point.col = "black",  # black points
                  bar.col = "gray"   # gray bars
                  )
```

# Element opacities

You can adjust the opacities of individual plotting elements in one of two ways. The theme.o argument allows you to specify an opacity theme. A value of 2 will turn off the inference boxes and turn up the bars:

```{r, fig.width = 6, fig.height = 6}
yarrr::pirateplot(formula = weight ~ Diet,
                  data = ChickWeight,
                  theme.o = 2, # Use opacity theme 2
                  main = "Opacity themes with theme.o"
                  )
```

A theme.o value of 0 will turn off all of the opacities. You can then manually turn on each one with the x.o arguments (e.g.; bean.o, point.o, line.o)

```{r, fig.width = 6, fig.height = 6}
yarrr::pirateplot(formula = weight ~ Diet,
                  data = ChickWeight,
                  theme.o = 0, # Turn off all opacities,
                  point.o = .1, # Turn point opacities up to .1
                  inf.o = .3, # Turn inf opacities up to .3
                  bean.o = .9, # Turn beans way up
                  main = "Changing opacities individually with point.o, inf.o (etc.)"
                  )
```

You can specify the x.o arguments as *vectors* to selectively turn up (or down) specific elements for specific groups:

```{r, fig.width = 6, fig.height = 6}
yarrr::pirateplot(formula = weight ~ Diet,
                  data = ChickWeight,
                  pal = "basel",
                  theme.o = 0, # Turn off all opacities,
                  point.o = c(.1, .05, .2, .05), # Selectively turn up point opacities
                  inf.o = c(.9, .1, .9, .1), # Selectively turn up inf opacities
                  bean.o = c(1, .1, 1, .1), # Selective bean opacity adjustment
                  main = "Adjusting opacities between groups"
                  )
```

# Additional arguments

You can use up to two independent variables in the formula, change the color of the background with back.col, and the color of the gridlines with gl.col:

```{r, fig.width = 10, fig.height = 6}
yarrr::pirateplot(formula = weight ~ Diet + Time,
                  data = ChickWeight,
                  pal = "google",
                  theme.o = 1, # Turn off all opacities,
                  back.col = gray(.96),
                  gl.col = "white",
                  main = "back.col, gl.col"
                  )
```

# Contact

For questions, comments, or suggestions, feel free to write me at yarrr.book@gmail.com


# References


